Sound Automated Synthesis of Digital Stabilizing Controllers for Continuous Plants

### 2016-10-13 paper/paper.tex
line breaks

%
In figure~\ref{fig:digsystem1}, two sources of uncertainties are considered: parametric uncertainties related to modelling errors, and uncertainties introduced by quantisations in sampling and holding process, {\it i.e.} ADC and DAC conversions. Assuming linearity and independence between these sources, the uncertain model may be rewritten as follows:
\begin{equation}
\label{eq:complete_unc_model}
\vec{\hat{G}}=\vec{G}+\Delta_p \vec{G}+\Delta_q \vec{G},
\end{equation}
%
where $\Delta_p \vec{G}$ is the term that corresponds to the parametric uncertainties and $\Delta_q \vec{G}$ is the term that reflects the quantisation effects on the plant model. %\red{[next figure is a bit unnecessary, does not ad much to the previous one.]} Figure~\ref{fig:digsystem2} shows the digital closed loop systems with $\Delta G_{P}(z)$ and $\Delta G_{Q}(z)$.

@@

Let $\hat{C}(z)$ be the transfer function of the digital controller represented using this FWL with integer size $I$ and fractional
size $F$.  We shall denote the FWL domain of coefficients as $\mathbb{R}\langle I,F \rangle$
%Since the controller is implemented in the $\mathbb{R}\langle I,F \rangle$ 
%domain, {\it i.e.} a Fixed-Point Finite Word Length with integer size $I$ and fractional
%size $F$, we shall denote this domain as $\mathbb{R}\langle I,F \rangle$ 
and assume that there exists a function $FWL[\cdot]:\mathcal{P}^{n}\rightarrow \mathcal{P}^{n}\langle I,F \rangle$, which applies the FWL effects to a polynomial. $\mathcal{P}^{n}$ is the space of polynomials of $n$-th order and $\mathcal{P}^{n}\langle I,F \rangle$ is is the space of polynomials with coefficients in $\mathbb{R}\langle I,F \rangle$.
%, \red{[explain index n! why do you need to work on a Euclidean domain?]}. 
Thus the perturbed controller model $\hat{C}(z)$ may be computed from the original model $C(z)=\frac{N_{C}(z)}{D_{C}(z)}$ using the following expression:
\begin{equation}
\hat{C}(z)=\frac{FWL[N_{C}(z)]}{FWL[D_{C}(z)]}.
\end{equation}
In the case of a synthesised controller, because the synthesis is performed using FWL, $\hat{C}(z) \equiv C(z)$.

@@

In order to obtain a sound synthesis of the digital controller, the effect of FWL implementation in controller and quantisation noises in closed loop system have to be considered. Considering that the quantizer $Q1$ is the source of a white noise $\nu_{1}$ and $Q2$ is the source of a white noise $\nu_{2}$, the following equation is written for the system in Figure~\ref{fig:sampledsystem}, considering the parametric uncertainties $\Delta \vec{G}$ and the FWL effects on controller ($\hat{C}(z)$):
\begin{equation}
\hat{Y}(z)=\hat{C}(z)\hat{G}(z)[R(z)+\nu_{1}(z)]+\nu_{2}(z)\hat{G}(z)-\hat{Y}(z)\hat{C}(z)\hat{G}(z),
\end{equation}
that can be rewritten as follows:
\begin{equation}
\label{eq:outputfunctions}
\hat{Y}(z)=H_{1}[R(z)+\nu_{1}(z)]+H_{2}(z)\nu_{2}(z),
\end{equation}
where
\begin{equation}
H_{1}=\frac{\hat{C}(z)\hat{G}(z)}{1+\hat{C}(z)\hat{G}(z)},
\end{equation}
and
\begin{equation}
H_{2}=\frac{\hat{G}(z)}{1+\hat{C}(z)\hat{G}(z)},
\end{equation}

@@

\begin{mylemma}
\label{lemma:variance}
In particular, Widrow \cite{widrow1956} models the quantisation noise as a gaussian
model in which the second moment (variance) of the probability density
function (PDF) of the quantisation noise is given by
%
\begin{equation}
\label{eq:variancelemma}
\sigma_{Q}^{2}=\frac{q^{2}}{12},
\end{equation}
where $q$ is the step size of quantizer. (Proof is on \cite{widrow1956})
\end{mylemma}

\begin{mylemma}
\label{lemma:boundednoise}
The power spectral density of the quantisation noise (supposed white) $S_{\nu}(z)$ is bounded and may be computed through the z-transform of autocorrelation function of $\nu(k)$, thus:
\begin{equation}
S_{\nu}(z)= \sigma_{Q}^{2}=\frac{q^{2}}{12}.
\end{equation}
\end{mylemma}

\begin{proof}
Considering $\nu(k)$ a random process, where $\nu(k)$ is independent from $\nu(k+i)$ for every  integer $k\neq 0$. The autocorrelation function of $\nu(k)$, {\it i.e.} $R_{\nu\nu}(\tau)$ may be computed as follows:
\begin{equation}
R_{\nu\nu}(\tau)=E[\nu(t+\tau)\nu(t)]=\sigma_{Q}^{2}\cdot \delta(\tau),
\end{equation}
where $\delta(\tau)$ is the Dirac delta function at delay time $\tau$.

By definition~\cite{poularikas2000transforms}, the power spectral density of  $\nu(k)$ is the z-transform of $R_{\nu\nu}(\tau)$, thus
\begin{equation}
S_{\nu}(z)=  \sum\limits_{k=-\infty}^{\infty}{R_{\nu\nu}(k)z^{-k}} = \sigma_{Q}^{2},
\end{equation}
considering the Lemma~\ref{lemma:variance}:
\begin{equation}
S_{\nu}(z)=  \frac{q^{2}}{12}.
\end{equation}
\end{proof}

@@

\begin{mylemma}{\cite{Bessa16,fadali}}
\label{lemma:eq_int_stab} 
A feedback hybrid control system as shown in Fig.~\ref{fig:sampledsystem}, with a FWL implementation of the digital controller $\hat{C}(z) = \frac{FWL[N_{C}(z)]}{FWL[D_{C}(z)]}$ and uncertain discrete model of the plant $\vec{\hat{G}}=\vec{G}+\Delta_p \vec{G}+\Delta_q \vec{G} \Rightarrow \hat{G}(z)=\frac{\hat{N}_{G}(z)}{\hat{D}_{G}(z)}$, is BIBO-stable if and only if: 
\begin{itemize}
\item  the roots of characteristic polynomial $S(z)$ are inside the open unit circle, where $S(z)$ is:
\begin{equation}
\label{eq:internal_stab_lemma}
S(z)=FWL[N_{C}(z)]\hat{N}_{G}(z)+FWL[D_{C}(z)]\hat{D}_{G}(z);
\end{equation}
\item the direct loop product $\frac{FWL[N_{C}(z)]}{FWL[D_{C}(z)]}\cdot \frac{\hat{N}_{G}(z)}{\hat{D}_{G}(z)}$ has no pole-zero cancellation on or outside the unit circle.
\end{itemize}
\end{mylemma}

@@

An overview of the synthesis process is illustrated in Fig.~\ref{DSVerifier_process};
a detailed description of DSVerifier tool can be found in~\cite{IsmailBCFF15}.
Steps $1$ to $5$ are performed by the user/synthesiser and
Steps A to D are automatically performed by DSVerifier.
%, which accepts the
%digital controller transfer functions together with the plant model.  For
%any digital controller, implementation details should be provided ({\it
%e.g.}, number of bits, realization, and sample time).

@@

In order to synthesise closed-loop digital control systems, we use a
program synthesis engine.  Nowadays, such engines are used
increasingly often in program
verification~\cite{DBLP:conf/lpar/DavidKL15,DBLP:conf/cav/0001A14}.
Our program synthesiser makes use of Counter-Example Guided Inductive
Synthesis (CEGIS) \cite{sketch}.  We start by presenting its general
architecture followed by describing the parts specific to closed-loop
control systems.

@@

The design of our synthesiser is given in Fig.~\ref{fig:CEGIS} and
consists of two phases, {\sc Synthesise} and {\sc Verify}, which
interact via a finite set of test vectors {\sc inputs} that is updated
incrementally.  Given the aforementioned specification $\sigma$, the
{\sc synth} procedure tries to find an existential witness $\vec{P}$
satisfying the specification $\sigma(\vec{x}, \vec{P})$ for all
$\vec{x}$ in {\sc inputs} (as opposed to all $\vec{x} \in \mathcal{D}$).
%
If {\sc synthesise} succeeds in finding an witness $\vec{P}$, this
witness is a candidate solution to the full synthesis formula. 
We pass this candidate solution to {\sc verify}, which 
checks whether it is a full solution ({\it i.e.}, $\vec{P}$
satisfies the specification $\sigma(\vec{x}, \vec{P})$ for all
$\vec{x}\in\mathcal{D}$).
%determines whether
%it does satisfy the specification on all inputs by checking
%satisfiability of the following verification formula:
%
%$\exists \vec{x} . \lnot \sigma(\vec{x}, \vec{P})$
%If this formula is unsatisfiable, 
%% the candidate solution is in fact a 
%% full solution.
If this is the case, then the algorithm terminates.
Otherwise, additional information is provided to the {\sc synthesise}
phase in the form of a new counterexample that is added to the {\sc inputs} set
and the loop iterates again (note that, for now, we ignore the 
second feedback signal ``Increase Precision'' provided by the {\sc Verify}
phase in Fig.~\ref{fig:CEGIS} as it is specific to control synthesis and will 
be described in the next section).

@@

Since the controller is synthesised in the $\mathbb{R}\langle I,F \rangle$
domain,
%
 $$\forall i \leq N_C, j \leq M_C\ \  \tilde \beta_{i} \equiv \beta_{i} \wedge \tilde \alpha_{j} \equiv \alpha_{j} \Leftrightarrow \tilde C(z) \equiv C(z)$$
%
However, given a real plant $\hat{G}(z)$, and a discretising function
\begin{align}
\small
\label{FWL_tf}
&f_b (x \in \mathbb{R}) \rightarrow \tilde x \in \mathbb{R}\langle I_b,F_b \rangle \triangleq \tilde x = x-(x\mod\tilde x) \\
&FWL_b[P \in \mathcal{P}] = \tilde P \in \mathcal{P}^{n}\langle I_b,F_b \rangle : c_i \in \vec{P} \wedge \tilde c_i \in \tilde \vec{P}=f_b(c_i)  \nonumber
\end{align}
we calculate $\tilde G(z)=FWL_b(\hat{G}(z))$
\begin{align}
% \label{digital_controller_tf}
% \tilde C(z)&=\frac{(\hat{\beta}_{0}+\Delta_q \hat{\beta}_{0}) +...+(\hat{\beta}_{M_C}+\Delta_b \hat{\beta}_{M_C})z^{-M_C}}{(\hat{\alpha}_{0}+\Delta_q \hat{\alpha}_{0})+...+(\hat{\alpha}_{N_C}+\Delta_q \hat{\alpha}_{N_C})z^{-N_G}} \nonumber \\
% &= C(z) =\hat{C}(z) + \Delta{C}_b(z)\\
\label{digital_plant_tf}
\tilde G(z)&=\frac{(\hat{b}_{0}+\Delta_b \hat{b}_{0}) +...+(\hat{b}_{M_G}+\Delta_b \hat{b}_{M_G})z^{-M_G}}{(\hat{a}_{0}+\Delta_b \hat{a}_{0})+...+(\hat{a}_{N_G}+\Delta_b \hat{a}_{N_G})z^{-N_G}} \nonumber \\
&\Rightarrow \vec{\tilde G} =\vec{\hat{G}}+\Delta_b{\vec{G}}=\vec{G}+\Delta_p{\vec{G}}+\Delta_b{\vec{G}}.
\end{align}
where $\Delta_b\hat{c}_i=\hat{c}_i\mod \tilde{c}_i$
and $\Delta_b{G}$ represents the plant uncertainty caused by
the rounding off effect.  We may encompass all the uncertainty as
$\Delta{\vec{G}}=\Delta_p{\vec{G}}+\Delta_b{\vec{G}}$.  % Since this value is
% nondeterministic, the set of possible polynomial coefficients of $\tilde G(z)$ is
% denoted by $\mathfrak{G}(z)$.

@@

This time the {\sc uncertainty} check doesn't find any more counterexamples 
and we pass the current candidate solution to the {\sc precision} verification stage.
% In the second {\sc verify} phase, we check whether the current candidate
% solution above is the final solution. It is not and the verifier signals
% that it is due to lack of precision.
We obtain the result $false$, meaning that the current precision is insufficient.
Consequently, we increase our precision to $I_b=20,F_b=28$.
%
Since the previous counterexamples were attained at lower precision, we 
remove them from the counterexample set. Back in the {\sc synthesise}
phase, we re-start the process with a candidate solution with all coefficients 0.
Next, the {\sc uncertainty} verification stage provides the first counterexample at higher
precision:
%
%$$
%\begin{array}{ll}
%\Delta N_G(z) {=} 0.026314\\
%\Delta D_G(z) {=} 0.999024z{-}1.004785
%\end{array}
%$$
$$
\tilde G(z) = \frac{0.026314}{0.999024z{-}1.004785}
$$

@@

A third set of benchmarks that consists of a physical plant for satellite
applications is considered~\cite{Franklin15}, in order to evaluate the
present approach full capabilities.  In particular, satellites often require
attitude (pose) control for proper orientation of antennas and sensors
w.r.t.  earth.  The satellite attitude control is typically used for
three-axis attitude tracking, but here only one axis at a time is
considered.  The motion equation for the one-axis system, despising
disturbance torques, is given by
%
\begin{equation}
\label{eq:satelliteode}
I\cdot \ddot{\theta} = \tau_{C},
\end{equation}

@@

The fourth (and last) set of benchmarks considers the following description for the plant 
$G_{4}(s)$, which is typically used for evaluating stability margins~\cite{bhattacharyya97,keel_Bhattacharyya_examples}
%
\begin{equation}
\label{exampleA}
G_{4}(s)=\frac{s-1}{s^{2}-s-2}
\end{equation}
%
\noindent where $G_{4}(z)$ is obtained using the ZOH discretisation defined 
in Eq.~\eqref{eq:pulsetf} with $3$ different sample times ($0.1$, $0.05$, and $0.03$ 
seconds).

@@

The problem of parametric control synthesis based on stability measures for
continuous Linear Time Invariant (LTI) Single Input-Single Output (SISO)
systems has been researched for several decades.  On a theoretical level it
is a solved problem~\cite{wonham1967pole} for which researches continuously
seek better results based on a number of properties in addition to
stability.  A vast range of pole placement techniques such as Moore's
algorithm for eigenstructure assignment~\cite{klein1977eigenvalue} or the
more recent Linear Quadratic Regulator (LQR)~\cite{bemporad2002explicit}
have been used in several studies with increasing degrees of success to
ensure stability.  The latter approach highlights the importance of
conserving energy during the control process, which results in lower running
costs.  Since real systems are subject to tolerance and noise as well as the
need for economy, more recent studies focus on the problems of achieving
robust stability with minimum
gain~\cite{schmid2014unified,konigorski2012pole}.  However, when extending
the problem to digital controller synthesis, many of these techniques lack
the ability to produce sound or stable results because they disregard the
effects of quantisation and rounding.  Recent papers on
implementations/synthesis of LTI digital
controllers~\cite{das2013lqr,ghosh2013fpga} focus on time discretisation,
failing to account for these error-inducing effects and can therefore result
in digitally unstable systems that have been proven to be robustly stable in
a continuous space.  This situation demands the inclusion of formal methods
to verify the effects of space discretisation on the implementation of the
controller.

>>>>

%
In figure~\ref{fig:digsystem1}, two sources of uncertainties are considered: parametric uncertainties related to modelling errors, and uncertainties introduced by quantisations in sampling and holding process, {\it i.e.} ADC and DAC conversions. Assuming linearity and independence between these sources, the uncertain model may be rewritten as follows:
\begin{equation}
\label{eq:complete_unc_model}
\vec{\hat{G}}=\vec{G}+\Delta_p \vec{G}+\Delta_q \vec{G},
\end{equation}
%
where $\Delta_p \vec{G}$ is the term that corresponds to the parametric
uncertainties and $\Delta_q \vec{G}$ is the term that reflects the
quantisation effects on the plant model.

%\red{[next figure is a bit unnecessary, does not ad much to the previous one.]} Figure~\ref{fig:digsystem2} shows the digital closed loop systems with $\Delta G_{P}(z)$ and $\Delta G_{Q}(z)$.

@@

Let $\hat{C}(z)$ be the transfer function of the digital controller
represented using this FWL with integer size $I$ and fractional size $F$. 
We shall denote the FWL domain of coefficients as $\mathbb{R}\langle I,F
\rangle$
%
%Since the controller is implemented in the $\mathbb{R}\langle I,F \rangle$ 
%domain, {\it i.e.} a Fixed-Point Finite Word Length with integer size $I$ and fractional
%size $F$, we shall denote this domain as $\mathbb{R}\langle I,F \rangle$ 
and assume that there exists a function $FWL[\cdot]:\mathcal{P}^{n}\rightarrow \mathcal{P}^{n}\langle I,F \rangle$, which applies the FWL effects to a polynomial. $\mathcal{P}^{n}$ is the space of polynomials of $n$-th order and $\mathcal{P}^{n}\langle I,F \rangle$ is is the space of polynomials with coefficients in $\mathbb{R}\langle I,F \rangle$.
%, \red{[explain index n! why do you need to work on a Euclidean domain?]}. 
%
Thus the perturbed controller model $\hat{C}(z)$ may be computed from the
original model $C(z)=\frac{N_{C}(z)}{D_{C}(z)}$ using the following
expression:
%
\begin{equation}
\hat{C}(z)=\frac{FWL[N_{C}(z)]}{FWL[D_{C}(z)]}.
\end{equation}
%
In the case of a synthesised controller, because the synthesis is performed
using FWL, $\hat{C}(z) \equiv C(z)$.

@@

In order to obtain a sound synthesis of the digital controller, the effect
of FWL implementation in controller and quantisation noises in closed loop
system have to be considered.  Considering that the quantizer $Q1$ is the
source of a white noise $\nu_{1}$ and $Q2$ is the source of a white noise
$\nu_{2}$, the following equation is written for the system in
Figure~\ref{fig:sampledsystem}, considering the parametric uncertainties
$\Delta \vec{G}$ and the FWL effects on controller ($\hat{C}(z)$):
%
\begin{equation}
\hat{Y}(z)=\hat{C}(z)\hat{G}(z)[R(z)+\nu_{1}(z)]+\nu_{2}(z)\hat{G}(z)-\hat{Y}(z)\hat{C}(z)\hat{G}(z),
\end{equation}
%
that can be rewritten as follows:
%
\begin{equation}
\label{eq:outputfunctions}
\hat{Y}(z)=H_{1}[R(z)+\nu_{1}(z)]+H_{2}(z)\nu_{2}(z),
\end{equation}
%
where
%
\begin{equation}
H_{1}=\frac{\hat{C}(z)\hat{G}(z)}{1+\hat{C}(z)\hat{G}(z)},
\end{equation}
%
and
%
\begin{equation}
H_{2}=\frac{\hat{G}(z)}{1+\hat{C}(z)\hat{G}(z)},
\end{equation}

@@

\begin{mylemma}
\label{lemma:variance}
%
In particular, Widrow \cite{widrow1956} models the quantisation noise as a
gaussian model in which the second moment (variance) of the probability
density function (PDF) of the quantisation noise is given by
%
\begin{equation}
\label{eq:variancelemma}
\sigma_{Q}^{2}=\frac{q^{2}}{12},
\end{equation}
where $q$ is the step size of quantizer. (Proof is on \cite{widrow1956})
\end{mylemma}

\begin{mylemma}
\label{lemma:boundednoise}
%
The power spectral density of the quantisation noise (supposed white)
$S_{\nu}(z)$ is bounded and may be computed through the z-transform of
autocorrelation function of $\nu(k)$, thus:
%
\begin{equation}
S_{\nu}(z)= \sigma_{Q}^{2}=\frac{q^{2}}{12}.
\end{equation}
\end{mylemma}

\begin{proof}
%
Considering $\nu(k)$ a random process, where $\nu(k)$ is independent from
$\nu(k+i)$ for every integer $k\neq 0$.  The autocorrelation function of
$\nu(k)$, {\it i.e.} $R_{\nu\nu}(\tau)$ may be computed as follows:
%
\begin{equation}
R_{\nu\nu}(\tau)=E[\nu(t+\tau)\nu(t)]=\sigma_{Q}^{2}\cdot \delta(\tau),
\end{equation}
%
where $\delta(\tau)$ is the Dirac delta function at delay time $\tau$.

By definition~\cite{poularikas2000transforms}, the power spectral density of
$\nu(k)$ is the z-transform of $R_{\nu\nu}(\tau)$, thus
%
\begin{equation}
S_{\nu}(z)=  \sum\limits_{k=-\infty}^{\infty}{R_{\nu\nu}(k)z^{-k}} = \sigma_{Q}^{2},
\end{equation}
considering the Lemma~\ref{lemma:variance}:
\begin{equation}
S_{\nu}(z)=  \frac{q^{2}}{12}.
\end{equation}
\end{proof}

@@

\begin{mylemma}{\cite{Bessa16,fadali}}
\label{lemma:eq_int_stab} 
%
A feedback hybrid control system as shown in Fig.~\ref{fig:sampledsystem},
with a FWL implementation of the digital controller $\hat{C}(z) =
\frac{FWL[N_{C}(z)]}{FWL[D_{C}(z)]}$ and uncertain discrete model of the
plant $\vec{\hat{G}}=\vec{G}+\Delta_p \vec{G}+\Delta_q \vec{G} \Rightarrow
\hat{G}(z)=\frac{\hat{N}_{G}(z)}{\hat{D}_{G}(z)}$, is BIBO-stable if and
only if:
%
\begin{itemize}
\item  the roots of characteristic polynomial $S(z)$ are inside the open unit circle, where $S(z)$ is:
\begin{equation}
\label{eq:internal_stab_lemma}
S(z)=FWL[N_{C}(z)]\hat{N}_{G}(z)+FWL[D_{C}(z)]\hat{D}_{G}(z);
\end{equation}
\item the direct loop product $\frac{FWL[N_{C}(z)]}{FWL[D_{C}(z)]}\cdot \frac{\hat{N}_{G}(z)}{\hat{D}_{G}(z)}$ has no pole-zero cancellation on or outside the unit circle.
\end{itemize}
\end{mylemma}

@@

An overview of the synthesis process is illustrated in
Fig.~\ref{DSVerifier_process}; a detailed description of DSVerifier tool can
be found in~\cite{IsmailBCFF15}.  Steps $1$ to $5$ are performed by the
user/synthesiser and Steps A to D are automatically performed by DSVerifier.
%
%, which accepts the
%digital controller transfer functions together with the plant model.  For
%any digital controller, implementation details should be provided ({\it
%e.g.}, number of bits, realization, and sample time).

@@

In order to synthesise closed-loop digital control systems, we use a program
synthesis engine.  Nowadays, such engines are used increasingly often in
program verification~\cite{DBLP:conf/lpar/DavidKL15, DBLP:conf/cav/0001A14}. 
Our program synthesiser makes use of Counter-Example Guided Inductive
Synthesis (CEGIS) \cite{sketch}.  We start by presenting its general
architecture followed by describing the parts specific to closed-loop
control systems.

@@

The design of our synthesiser is given in Fig.~\ref{fig:CEGIS} and consists
of two phases, {\sc Synthesise} and {\sc Verify}, which interact via a
finite set of test vectors {\sc inputs} that is updated incrementally. 
Given the aforementioned specification $\sigma$, the {\sc synth} procedure
tries to find an existential witness $\vec{P}$ satisfying the specification
$\sigma(\vec{x}, \vec{P})$ for all $\vec{x}$ in {\sc inputs} (as opposed to
all $\vec{x} \in \mathcal{D}$).
%
If {\sc synthesise} succeeds in finding an witness $\vec{P}$, this witness
is a candidate solution to the full synthesis formula.  We pass this
candidate solution to {\sc verify}, which checks whether it is a full
solution ({\it i.e.}, $\vec{P}$ satisfies the specification $\sigma(\vec{x},
\vec{P})$ for all $\vec{x}\in\mathcal{D}$).
%
%determines whether
%it does satisfy the specification on all inputs by checking
%satisfiability of the following verification formula:
%
%$\exists \vec{x} . \lnot \sigma(\vec{x}, \vec{P})$
%If this formula is unsatisfiable, 
%% the candidate solution is in fact a 
%% full solution.
%
If this is the case, then the algorithm terminates.  Otherwise, additional
information is provided to the {\sc synthesise} phase in the form of a new
counterexample that is added to the {\sc inputs} set and the loop iterates
again (note that, for now, we ignore the second feedback signal ``Increase
Precision'' provided by the {\sc Verify} phase in Fig.~\ref{fig:CEGIS} as it
is specific to control synthesis and will be described in the next section).

@@

Since the controller is synthesised in the $\mathbb{R}\langle I,F \rangle$
domain,
%
 $$\forall i \leq N_C, j \leq M_C\ \  \tilde \beta_{i} \equiv \beta_{i} \wedge \tilde \alpha_{j} \equiv \alpha_{j} \Leftrightarrow \tilde C(z) \equiv C(z)$$
%
However, given a real plant $\hat{G}(z)$, and a discretising function
%
\begin{align}
\small
\label{FWL_tf}
&f_b (x \in \mathbb{R}) \rightarrow \tilde x \in \mathbb{R}\langle I_b,F_b \rangle \triangleq \tilde x = x-(x\mod\tilde x) \\
&FWL_b[P \in \mathcal{P}] = \tilde P \in \mathcal{P}^{n}\langle I_b,F_b \rangle : c_i \in \vec{P} \wedge \tilde c_i \in \tilde \vec{P}=f_b(c_i)  \nonumber
\end{align}
%
we calculate $\tilde G(z)=FWL_b(\hat{G}(z))$
%
\begin{align}
% \label{digital_controller_tf}
% \tilde C(z)&=\frac{(\hat{\beta}_{0}+\Delta_q \hat{\beta}_{0}) +...+(\hat{\beta}_{M_C}+\Delta_b \hat{\beta}_{M_C})z^{-M_C}}{(\hat{\alpha}_{0}+\Delta_q \hat{\alpha}_{0})+...+(\hat{\alpha}_{N_C}+\Delta_q \hat{\alpha}_{N_C})z^{-N_G}} \nonumber \\
% &= C(z) =\hat{C}(z) + \Delta{C}_b(z)\\
\label{digital_plant_tf}
\tilde G(z)&=\frac{(\hat{b}_{0}+\Delta_b \hat{b}_{0}) +...+(\hat{b}_{M_G}+\Delta_b \hat{b}_{M_G})z^{-M_G}}{(\hat{a}_{0}+\Delta_b \hat{a}_{0})+...+(\hat{a}_{N_G}+\Delta_b \hat{a}_{N_G})z^{-N_G}} \nonumber \\
&\Rightarrow \vec{\tilde G} =\vec{\hat{G}}+\Delta_b{\vec{G}}=\vec{G}+\Delta_p{\vec{G}}+\Delta_b{\vec{G}}.
\end{align}
where $\Delta_b\hat{c}_i=\hat{c}_i\mod \tilde{c}_i$
and $\Delta_b{G}$ represents the plant uncertainty caused by
the rounding off effect.  We may encompass all the uncertainty as
$\Delta{\vec{G}}=\Delta_p{\vec{G}}+\Delta_b{\vec{G}}$.  % Since this value is
% nondeterministic, the set of possible polynomial coefficients of $\tilde G(z)$ is
% denoted by $\mathfrak{G}(z)$.

@@

This time the {\sc uncertainty} check doesn't find any more counterexamples 
and we pass the current candidate solution to the {\sc precision} verification stage.
% In the second {\sc verify} phase, we check whether the current candidate
% solution above is the final solution. It is not and the verifier signals
% that it is due to lack of precision.
We obtain the result $false$, meaning that the current precision is insufficient.
Consequently, we increase our precision to $I_b=20,F_b=28$.
%
Since the previous counterexamples were attained at lower precision, we
remove them from the counterexample set.  Back in the {\sc synthesise}
phase, we re-start the process with a candidate solution with all
coefficients $0$.  Next, the {\sc uncertainty} verification stage provides
the first counterexample at higher precision:
%
%$$
%\begin{array}{ll}
%\Delta N_G(z) {=} 0.026314\\
%\Delta D_G(z) {=} 0.999024z{-}1.004785
%\end{array}
%$$
$$
\tilde G(z) = \frac{0.026314}{0.999024z{-}1.004785}
$$

@@

A third set of benchmarks that consists of a physical plant for satellite
applications is considered~\cite{Franklin15}, in order to evaluate the
present approach full capabilities.  In particular, satellites often require
attitude (pose) control for proper orientation of antennas and sensors
w.r.t.~earth.  The satellite attitude control is typically used for
three-axis attitude tracking, but here only one axis at a time is
considered.  The motion equation for the one-axis system, despising
disturbance torques, is given by
%
\begin{equation}
\label{eq:satelliteode}
I\cdot \ddot{\theta} = \tau_{C},
\end{equation}

@@

The fourth (and last) set of benchmarks considers the following description
for the plant $G_{4}(s)$, which is typically used for evaluating stability
margins~\cite{bhattacharyya97, keel_Bhattacharyya_examples}
%
\begin{equation}
\label{exampleA}
G_{4}(s)=\frac{s-1}{s^{2}-s-2}
\end{equation}
%
\noindent where $G_{4}(z)$ is obtained using the ZOH discretisation defined 
in Eq.~\eqref{eq:pulsetf} with $3$ different sample times ($0.1$, $0.05$, and $0.03$ 
seconds).

@@

The problem of parametric control synthesis based on stability measures for
continuous Linear Time Invariant (LTI) Single Input-Single Output (SISO)
systems has been researched for several decades.  On a theoretical level it
is a solved problem~\cite{wonham1967pole} for which researches continuously
seek better results based on a number of properties in addition to
stability.  A vast range of pole placement techniques such as Moore's
algorithm for eigenstructure assignment~\cite{klein1977eigenvalue} or the
more recent Linear Quadratic Regulator (LQR)~\cite{bemporad2002explicit}
have been used in several studies with increasing degrees of success to
ensure stability.  The latter approach highlights the importance of
conserving energy during the control process, which results in lower running
costs.  Since real systems are subject to tolerance and noise as well as the
need for economy, more recent studies focus on the problems of achieving
robust stability with minimum gain~\cite{schmid2014unified,
konigorski2012pole}.  However, when extending the problem to digital
controller synthesis, many of these techniques lack the ability to produce
sound or stable results because they disregard the effects of quantisation
and rounding.  Recent papers on implementations/synthesis of LTI digital
controllers~\cite{das2013lqr,ghosh2013fpga} focus on time discretisation,
failing to account for these error-inducing effects and can therefore result
in digitally unstable systems that have been proven to be robustly stable in
a continuous space.  This situation demands the inclusion of formal methods
to verify the effects of space discretisation on the implementation of the
controller.
